The impact of cosmic variance on simulating weak lensing surveys

### 2014-12-01 Writeup/paper_v3.tex
type-dependent stuff

@@

\rachel{The estimates in this subsection are rough illustrations of
  the magnitudes of these effects.  Other aspects of shear calibration
  that could be affected relate to the use of {\em HST} data to
  estimate the impact of detailed galaxy morphology, or to calibrate
  the effect of colour gradients
  \citep{2012MNRAS.421.1385V,2013MNRAS.432.2385S}.  In the latter
  case, what is most relevant in this work is our finding that $B/T$
  exhibits environmental dependence, which likely translates into
  environmental dependence of colour gradients.  Unfortunately, we
  cannot directly test the strength of any colour gradient variations
  with environment using COSMOS data, due to the fact that there is
  only single-band coverage in much of its area.  We note that our
  findings may seem to be at odds with the conclusions in
  \cite{2013MNRAS.432.2385S} based on synthetic galaxy models that the
  existing area of {\em HST} coverage with $\ge 2$ bands is sufficient
  for calibration of colour
  gradients\footnote{\rachel{\cite{2012MNRAS.421.1385V} showed that the
    fluctuations in colour gradients within the source galaxy sample
    in the Euclid weak lensing survey due to environmental effects
    will not be a significant source of uncertainty.  However, the
    question addressed in that work is different from the question
    considered here, which is the impact of fluctuations in colour
    gradients in the training sample used to derive the corrections,
    rather than in the galaxy sample to which those corrections will
    be applied.}}.  This is particularly striking given that, as shown
  there, the dominant galaxy sample with $2$-band coverage is the
  AEGIS dataset, which has substantially smaller area than COSMOS, and
  therefore should exhibit a stronger influence of cosmic variance.
  \rachel{The presence of other fields besides AEGIS (e.g., GOODS)
    should help mitigate this effect given that the large-scale
    structure in the two fields will be completely uncorrelated, but
    the number of fields is still not large and the area is dominated
    by a small number of them\footnote{\rachel{For context,
        \cite{2015APh....63...81N} 
        showed that for tests of photometric redshift quality, the
        number of spectroscopic fields of this size that would be
        required for future surveys to avoid the influence of cosmic
        variance ranges from many tens to hundreds}}.}  It seems that a future study using
  an {\em HST} field with more than one band would be warranted, to
  test whether the discrepancy between our findings and those of
  \cite{2013MNRAS.432.2385S} arises from a weak connection between
  colour gradients and $B/T$ (meaning that our results do not imply a
  problem with colour gradient calibration), or from some unrealistic
  aspect to the synthetic models used in that work.}

@@

\rachel{The authors would like to thank Thomas Kitching and Henk
  Hoekstra for their helpful comments on this work.}
AK and RM acknowledge the support of NASA ROSES 12-EUCLID12-0004, and
program HST-AR-12857.01-A, provided by NASA through a grant from the
Space Telescope Science Institute, which is operated by the
Association of Universities for Research in Astronomy, Incorporated,
under NASA contract NAS5-26555. RM acknowledges the support of an Alfred P. Sloan Research Fellowship. 
Kavli IPMU is supported by World Premier International Research Center Initiative (WPI), MEXT, Japan.
We thank Alexie Leauthaud for 
many useful discussions.

>>>>

\rachel{As a test of the second option, we consider whether we can use some division of the
galaxies by color or $B/T$ (as a proxy for color) in order to
determine the structural parameters of galaxies as a function of
redshift for each of them separately.  If the effects we are seeing
can be cleanly described in terms of overdense (underdense) regions
being dominated by bulge- (disk-)dominated galaxies, then division
into samples based on $B/T$ may help reduce the effects that we have
described in previous subsections.  We test this idea by dividing the
sample into bulge- and disk-dominated galaxies based on requiring
$B/T\ge $ or $<0.4$.  The resulting bulge fraction $f_\text{bulge}$,
shown as a function of redshift in Fig.~\ref{fig:fbulge_z}, 
shows a mild effect from the large-scale structures in the COSMOS field.
}
\begin{figure}
 \centering
 \includegraphics[width=1.0\columnwidth]{fbulge_z} 
 \caption{\rachel{The fraction of galaxies in \s1 (open symbols) and
     \s3 (solid symbols) that are classified as bulge-dominated as a
     function of redshift, including the correlation with redshift
     slices classified as overdense or underdense as indicated by the
     symbol shape and color.}
         }
 \label{fig:fbulge_z}
\end{figure}

\rachel{Fig.~\ref{fig:rmsellip_bulge_disk} shows the RMS distortion as
  a function of redshift for the full \s1 sample (as shown previously) and
  for the separate bulge- and disk-dominated samples.  As shown,
  the deviations in the values due to large-scale structure are
  clearly evident in all three cases, though slightly more prominent
  for the full sample and less so for the bulge-dominated sample.  For
comparison with the results for the full sample shown in
Table~\ref{table:pvalues_all}, we compute the KS (AD) test $p$-values
for consistency of axis ratios between overdense and underdense
regions for bulge- and disk-dominated samples.  For \s1, these
$p$-values are 0.16 (0.08) for bulge-dominated galaxies and 0.005
(0.001) for disk-dominated galaxies.  For \s3, these $p$-values are
0.03 (0.03) and $10^{-4}$ ($3\times 10^{-4}$), respectively.  With the
possible exception of bulge-dominated galaxies, it seems that
separation into two morphological samples is not enough to remove the
effect of the local environment on the ensemble properties like the
intrinsic ellipticity distribution.   Due to the limited sample size,
we did not explore the possibility of an even finer morphological
division of the sample, but with deeper {\em HST} datasets, this may
be a promising path to pursue.  However, if the nature of the bulges
or disks themselves is altered by the local environment, then division
purely into bulge- or disk-dominated sample will not be helpful in
mitigating environmental effects.
}
\begin{figure}
 \centering
 \includegraphics[width=1.0\columnwidth]{rms_ellip_momentbased_S1_bulgedisk} 
 \caption{\rachel{The RMS distortion as a function of redshift for the
     full volume-limited galaxy sample \s3 (partially-full symbols), the
     disk-dominated galaxy sample (open symbols), and the
     bulge-dominated galaxy sample (full symbols).  The fact that the
     bulge-dominated sample has a lower RMS distortion than the
     disk-dominated sample is consistent with expectations.}
         }
 \label{fig:rmsellip_bulge_disk}
\end{figure}

@@

\rachel{The estimates in this subsection are rough illustrations of
  the magnitudes of these effects.  Other aspects of shear calibration
  that could be affected relate to the use of {\em HST} data to
  estimate the impact of detailed galaxy morphology, or to calibrate
  the effect of colour gradients
  \citep{2012MNRAS.421.1385V,2013MNRAS.432.2385S}.  In the latter
  case, what is most relevant in this work is our finding that $B/T$
  exhibits environmental dependence, which likely translates into
  environmental dependence of colour gradients.  Unfortunately, we
  cannot directly test the strength of any colour gradient variations
  with environment using COSMOS data, due to the fact that there is
  only single-band coverage in much of its area.  We note that our
  findings may seem to be at odds with the conclusions in
  \cite{2013MNRAS.432.2385S} based on synthetic galaxy models that the
  existing area of {\em HST} coverage with $\ge 2$ bands is sufficient
  for calibration of colour
  gradients as a function of galaxy type\footnote{\rachel{\cite{2012MNRAS.421.1385V} showed that the
    fluctuations in colour gradients within the source galaxy sample
    in the Euclid weak lensing survey due to environmental effects
    will not be a significant source of uncertainty.  However, the
    question addressed in that work is different from the question
    considered here, which is the impact of fluctuations in colour
    gradients in the training sample used to derive the corrections,
    rather than in the galaxy sample to which those corrections will
    be applied.}}.  This is particularly striking given that, as shown
  there, the dominant galaxy sample with $2$-band coverage is the
  AEGIS dataset, which has substantially smaller area than COSMOS, and
  therefore should exhibit a stronger influence of cosmic variance.
  \rachel{The presence of other fields besides AEGIS (e.g., GOODS)
    should help mitigate this effect given that the large-scale
    structure in the two fields will be completely uncorrelated, but
    the number of fields is still not large and the area is dominated
    by a small number of them\footnote{\rachel{For context,
        \cite{2015APh....63...81N} 
        showed that for tests of photometric redshift quality, the
        number of spectroscopic fields of this size that would be
        required for future surveys to avoid the influence of cosmic
        variance ranges from many tens to hundreds}}.}  It seems that a future study using
  an {\em HST} field with more than one band would be warranted, to
  test whether the discrepancy between our findings and those of
  \cite{2013MNRAS.432.2385S} arises from a weak connection between
  colour gradients and $B/T$ (meaning that our results do not imply a
  problem with colour gradient calibration), or from some unrealistic
  aspect to the synthetic models used in that work.}

@@

\rachel{The authors would like to thank Thomas Kitching, Henk
  Hoekstra, and Tim Schrabback for their helpful comments on this work.}
AK and RM acknowledge the support of NASA ROSES 12-EUCLID12-0004, and
program HST-AR-12857.01-A, provided by NASA through a grant from the
Space Telescope Science Institute, which is operated by the
Association of Universities for Research in Astronomy, Incorporated,
under NASA contract NAS5-26555. RM acknowledges the support of an Alfred P. Sloan Research Fellowship. 
Kavli IPMU is supported by World Premier International Research Center Initiative (WPI), MEXT, Japan.
We thank Alexie Leauthaud for 
many useful discussions.
